Unifying quasars with clumpy wind models

### 2015-11-10 compare.tex
updated figure

%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
\usepackage{url}
%\usepackage{pngfig}
%\usepackage[dvips]{color}mak
%\usepackage{aabib}

@@

%% Journal definitions
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\apjs{ApJs}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}
\def\aaps{A\&As}
\def\LA{Lyman\thinspace$\alpha$}

\newcommand{\EXPN}[2]{\mbox{$#1\times 10^{#2}$}}
\newcommand{\EXPU}[3]{\mbox{\rm $#1 \times 10^{#2} \rm\:#3$}}  % exponent with units
\newcommand{\POW}[2]{\mbox{$\rm10^{#1}\rm\:#2$}}
\def\LUM{\:{\rm ergs\:s^{-1}}}
\def\FLUX{\:{\rm ergs\:cm^{-2}\:s^{-1}}}
\def\OIGS{\:{\rm ergs\:cm^{-2}\:s^{-1}\:\AA^{-1}}}

@@

\title
{
% Modelling quasar outflows:
% clumpy winds and broad emission lines
Testing Quasar Unification with Clumpy Wind Models
%Unifying quasars with clumpy wind models
}

%\author[Matthews et al.]
\author{
\parbox[t]{\textwidth}{
James~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, Christian~Knigge$^1$,
Nick~Higginbottom$^1$, Knox~S.~Long$^2$, Stuart~A.~Sim$^3$ and Sam~W.~Mangham$^1$
}
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK
}

@@

Various unification schemes have been proposed to interpret the complex phenomenology 
of quasars and luminous active galactic nuclei (AGN) in terms of  a simple axisymmetric 
picture involving a central black hole, an accretion disc and an associated outflow.
Here, we continue our tests of  this paradigm by comparing ``average'' AGN spectra to 
synthetic spectra of simple biconical disc wind models, using our state-of-the-art 
Monte Carlo radiative transfer and photoionization code.
In particular, we investigate whether clumping of the outflow allows us to 
produce synthetic spectra in the rest-frame UV that have the characteristics of AGN, 
assuming total luminosities as high as 10$^{45}$ ergs s$^{-1}$. 

@@

allows for a more realistic AGN luminosities, while maintaining the ionization 
state necessary for strong BAL features in the rest-frame UV. 
We examine the X-ray properties of these simple clumped models

@@

The dense, X-ray heated wind produces strong recombination and collisionally 
excited line emission in, e.g., \civ\ and \la, to emerge at the low inclination, 
`Type 1 quasar-like' angles. 
At the highest inclinations, the synthetic spectra possess prominent \mg\ and \al\ BALs, the absorption features seen in LoBAL quasars. 
Despite these successes, we are unable to reproduce 

@@

@@

% {\bf This sentence needs to be more precise, especially in 
% light of the statement earlier in the abstract about ``good agreement''}

@@

@@

% {\bf There is confusion in the abstract, at least, in what is the model and what is the synthetic spectra, as well as whether there is one model or many models.  We need to try to be clear.  The model should be a clumped biconical outflow.  There are then particular instantiations of the model. And then there are synthetic spectra}

@@

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is often attributed to emission from a 
geometrically thin, optically thick accretion disc surrounding the central black hole, similar
to that described by \cite{shakurasunyaev1973}.
% {\bf Note: the first half of this paragraph and the second do not follow from one another.  
% The first half is about spectra the second about outflows and the connection to spectra is not made until the next paragrph.  
% This needs fixing. JM: Knox, I like this structure...I start with inflow, then move to outflow.}
In addition to the {\em inflowing} accreting material, 
{\em outflows} are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution \citep{springel2005}
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

Perhaps the clearest evidence of outflows in AGN is  
the blueshifted ($\sim 0.1c$) broad absorption lines (BALs) in the 
ultraviolet seen in approximately $20\%$ of quasars
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind. 
According to this paradigm, a biconical wind rises from 
the accretion disc and the BALQSO fraction is associated with
the covering factor of the outflow. 
Polarisation studies expect the wind to be roughly equatorial
\citep{goodrich1995, cohen1995}, although there is also evidence
for polar BAL outflows  in radio-loud (RL) sources \citep{zhou2006}.
% although this geometry is contradicted
% by more recent studies of radio spectral indices \citep{dipompeo2012}.

Due to their ubiquitous nature,
disc winds offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.
Within this unification framework, the broad-line region (BLR) can 
correspond either to the dense wind base or clumps embedded
in the outflow. Indeed, \cite{elitzur2014} show that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit high ionization line absorption; LoBALQSOs, which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs, which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified geometric models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

% As well as acting as a source of photoionized plasma, a 
% As well as imprinting clear line absorption and emission
% features, disc winds may also have a profound effect on the structure and 
% emergent {\em continuum} of the accretion disc itself.
% Mass-loss will alter the accretion rate and resultant 
% temperature of the accretion disc, possibly explaining some 
% of the features typically seen in luminous AGN \citep{knigge1999,laordavis2014}.
% There have been numerous difficulties when confronting 
% theoretical accretion disc models with observations 
% \citep[see e.g.][]{blaes1998}. However, AGN spectral energy distributions (SEDs) can now, 
% in general, be fitted well with accretion disc models
% when the effects of general relativity (GR), Comptonisation
% and mass-loss are included \citep{capellupo2015}. Mass-loss therefore appears to be 
% critical if an accretion disc model is to successfully fit AGN SEDs, 
% particular in the UV region of the spectrum.

%{\bf ksl: I don't see what this paragraph has to do with our problem, which we need to clearly explain} Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics,
Despite the clear importance of disc winds in shaping quasar and AGN spectra,  
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
% The presence of line-locked features \citep{bowler2014} 
% and the `ghost of \la' (Arav et al. 1996; Arav 1996; North 2006; but see 
% also Cottis et al. 2010) \nocite{arav1995, arav1996, north2006,cottis2010}
% in the spectra of some BALQSOs also gives clearer evidence that line-driving is
% at least partially contributing to the acceleration of the wind.
The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma, meaning that it is difficult to prevent 
the wind becoming over-ionized and `failing' in the presence of strong X-rays. 
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
% Hydrodynamic simulations of line-driven disc winds also found a shielding region
% was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
% However, \cite{sim2010_hydro} and \cite{H14} showed that including multiple scattering means the ionizing radiation field could still reach the previously shielded regions 
% in those particular models.
An additional or alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \citep{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required to explain the electron scattering wings of emission lines formed
in line-driven hot star winds \citep{hillier1991eswingsmodel}. Complex substructures 
are also produced in simulations of line-driven 
outflows in AGN, although on very different scales to line-driven instabilities 
\citep{PSK2000,PK04,progakurosawa2010,proga2014}.
Nevertheless, clumpy winds offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

@@

We have been engaged in a project to determine whether it is possible to simulate the properties of the spectra of AGN, including BAL QSOs, using simple kinematic prescriptions for biconical disc winds using a Monte Carlo radiative transfer (MCRT) code that calculates the ionization structure of the wind and simulates the spectra from such a system \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.  The results have encouraging in the sense that in H13, we showed we could produce simulated spectra that resembled that of BAL QSOs, as long as the luminosity of the central engine was relatively low, of order \POW{43}{\LUM} and the mass loss rate was relatively high, of order the mass accretion rate.  However, at higher luminosities, the wind was so ionized that UV absorption lines were not produced.  In addition, and in part due to limitations in our radiative transfer code, the model failed to produce spectra with strong emission lines at any inclination angle.

Here we attempt to address both of these issues, by introducing clumping into our model and a more complete treatment of H and He into our radiative transfer calculations.   The remainder of this paper is organized as follows:
% There has been some success using simple kinematic prescriptions for biconical disc winds to model  AGN and quasar outflows \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.  H13 successfully produced a benchmark model for BALQSOs. However, the model had two key drawbacks. First, an unrealistically low X-ray luminosity was required in order to prevent over-ionization of the outflow. Second, the model failed to produce the strong emission lines  required at low inclinations in a unified model. In this paper, we attempt to address these issues, and test the disc wind  unification model using Monte Carlo radiative transfer (MCRT) and photoionization  calculations. The paper is organised as follows.
In section 2, we describe some of the important photoionization 
and MCRT aspects of the code. We then outline the model in section 3, including 
a description of our clumping implementation and success criteria. 
Section 4 contains the results from a clumped model, which we discuss results
in comparison to observational data. Finally, we summarise our findings in section 5.

@@

For this study, we use the MCRT code \py\ we have developed to carry out our 
radiative transfer and photoionization simulations in non-local thermodynamic
equilibrium (non-LTE). The code can be used to model a variety of
disc-wind systems; it has been used with application to accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects 
\citep{simmacro2005} and quasars/AGN (H13, H14).\nocite{noebauer, M15, LK02}

The basic premise involves generating photons from an accretion disc
and central object, and computing their radiative transfer through an outflow.
This outflow is discretized into $n_x \times n_z$ cells in a 2.5D
cylindrical geometry with azimuthal symmetry.
First, the code gradually converges on the global temperature and ionization structure.
Once these ``ionization cycles'' are complete, then a series of further iterations
allow one to obtain a detailed synthetic spectrum.
LK02 provides the basic description of the original code; improvements, 
see especially \cite{simmacro2005}, H13 and M15, have been 
described in subsequent reports.  
%A full description of the code will be published elsewhere; 
We focus here on the specific changes made to improve the ionization calculation 
and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

% {\bf ksl - I a a bit torn over whether we add a few 3 sentences to describe the basic idea.  Some will be reading this for the first time.  Items that one would include are that we calculate the radiative transfer in a cylindrical coordinate system of dimension ?? x ??.  We have ionization and spectral cycles. In each cycle, ?? photon bundles are generated. We stop the ionization cycles after they converge, without any details.  I think this would be useful because we introduce terms like ``cell'' later on.', and without some kind of description they have no context}

@@

Our approach to line transfer is based upon the macro-atom implementation developed by \cite{lucy2002, lucy2003}. In our case, for reasons of computational effieciency, we adopt the  hybrid macro-atom scheme 
described by M15 
%(an improvement over that used by H14 for our earlier quasar study), 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interact with either two-level `simple ions'
or full `macro-atoms'. 
Our scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H is treated as a macro-atom, because 
we expect recombination to be important
in determining their level populations and resultant line emission, and because we were especially interested in the contribution to AGN spectra of \LA.  In H13, all atoms were treated in a two-level approximation.

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by Lucy (2002,2003). Previously (LK02, H13, M15),
we used a modified Saha approach to calculate the ionization fractions
of simple-ions. As part of  this effort, we have now improved {\sc PYTHON} to explicitly solve the 
rate equations between ions in non-LTE. This dispenses with a number of small assumptions 
made in the modified Saha approach, is more numerically stable, and, in principle, allows 
the direct addition of extra physical processes that would previously have necessitated 
approximate treatments.

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section, $n_e$
is the electron density and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative and 
collisional (three-body) recombination, respectively. 
$T_e$ is the electron temperature.
%% We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$. 
We stress that this approximation is not required for ions
treated as macro-atoms.

\subsection{Physical Processes}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% MODEL DESCRIPTION
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Our kinematic prescription for a biconical disc wind model
follows \cite{SV93}, and is described further by
LK02, H13 and M15. A schematic is shown in figure~\ref{fig:cartoon},
with key aspects marked. The general biconical
geometry is similar to that invoked by \cite{MCGV95} and 
\cite{elvis2000} to explain the phenomenonology
of quasars and BALQSOs.

@@

@@

The accretion disc in our model is geometrically thin, but optically thick.
We thus treat the disc as an ensemble of blackbodies with a 
\cite{shakurasunyaev1973} effective temperature profile. 
The emergent SED is then determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are assumed to be opaque, meaning
that photons which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~r_G$, where 
$r_G = GM_{BH}/c^2$ is the gravitational radius.
For a $10^9~M_\odot$ black hole, this is equal to $8.8\times10^{14}~{\rm cm}$ 
or $\sim10^{-4}~{\rm pc}$.

>>>>

%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
\usepackage{url}
%\usepackage{pngfig}
%\usepackage[dvips]{color}
%\usepackage{aabib}

@@

%% Journal definitions
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\apjs{ApJs}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}
\def\aaps{A\&As}

@@

\title
{
% Modelling quasar outflows:
% clumpy winds and broad emission lines
Unifying quasars with clumpy wind models
}

% [Matthews, J.]
% \author[Matthews, J.]{James Matthews \\
% {\sl  Supervisor: Prof. Christian Knigge} \\
% {\sl School of Physics \& Astronomy, University of Southampton,
%   Southampton, SO17 1BJ, UK}}

%\author[Matthews et al.]
\author{
\parbox[t]{\textwidth}{
J.~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, C.~Knigge$^1$,
N.~Higginbottom$^1$, K.~S.~Long$^2$, S.~A.~Sim$^3$ and S.~W.~Mangham$^1$
}
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK
}

@@

Various unification schemes for quasars and luminous active galactic nuclei (AGN) 
propose that much of their complex phenomenology can be explained by a simple 
geometrical picture involving an accretion disc and associated outflow.
Here, we test this paradigm by utilising our state-of-the-art radiative transfer 
and photoionization code to produce synthetic spectra from simple biconical disc wind models. 
In particular, we expand on our previous work by presenting a model that
exhibits many of the spectral features expected from a quasar unification model.

@@

allows for a more realistic X-ray luminosity in the model,
while maintaining the ionization state necessary for strong BAL features. 
We examine the X-ray properties of this new model

@@

We find that the dense, X-ray heated wind 
produces strong recombination and collisionally excited line emission in, 
e.g., \civ\ and \la, to emerge at the low inclination, `Type 1 quasar-like' angles. 
The highest inclination models possess prominent \mg\ and \al\ BALs, the absorption 
features seen in LoBAL quasars. Despite these successes, we are unable to reproduce 

@@

% remarkably similar of line-to-continuum ratios apparent 

@@

@@

We briefly explore the impact of general relativistic effects
on the angular distribution of disc continuum emission, and find that 
they are not sufficient to produce a self-consistent model.
% relativistic beaming, gravitational redshift and 
% We find that these general relativistic effects are not yet sufficient to produce a
% self-consistent model. 

@@

@@

% Introduction focussing on key points
% \begin{itemize}
% \item standard wind + BALQSO introduction
% \item focus on unification and that we are testing it
% \item some discussion of scales, referencing e.g. reverb maps, variability, microlensing, Arav 
% \item clumping background: stellar winds, clumping in AGN winds, variability
% \end{itemize}

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is often attributed to emssion from a 
geometrically thin, optically thick accretion disc surrounding the central black hole, similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material,
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution \citep{springel2005}
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

Approximately $20\%$ of quasars exhibit blueshifted ($\sim 0.1c$)
broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within the ADW paradigm, a biconical wind rises from 
the accretion disc and the BALQSO fraction is associated with
the covering factor of the outflow. 
The wind is expected to be roughly equatorial from polarisation results
\citep{goodrich1995, cohen1995}.
ADWs offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) can 
correspond either to the dense wind base or dense clumps embedded
in the outflows. Indeed, \cite{elitzur2014} found that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit higher ionization line absorption; LoBALQSOs which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

% As well as acting as a source of photoionized plasma, a 
As well as imprinting clear line absorption and emission
features, disc winds may also have a profound effect on the structure and 
emergent {\em continuum} of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features typically seen in luminous AGN \citep{knigge1999,laordavis2014}.
There have been numerous difficulties when confronting 
theoretical accretion disc models with observations 
\citep[see e.g.][]{blaes1998}. However, AGN spectral energy distributions (SEDs) can now, 
in general, be fitted well with accretion disc models
when the effects of general relativity (GR), Comptonisation
and mass-loss are included \citep{capellupo2015}. Mass-loss therefore appears to be 
critical if an accretion disc model is to successfully fit AGN SEDs, 
particular in the UV region of the spectrum.

Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics, 
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms for ADWs have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
The presence of line-locked features \citep{bowler2014} 
and the `ghost of \la' (Arav et al. 1996; Arav 1996; North 2006; but see 
also Cottis et al. 2010) \nocite{arav1995, arav1996, north2006,cottis2010}
in the spectra of some BALQSOs also gives clearer evidence that line-driving is
at least partially contributing to the acceleration of the wind.

The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma, meaning that it is difficult to prevent 
the wind becoming over-ionized and `failing' in the presence of strong X-rays. 
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.
An additional or alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \cite{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required to explain the electron scattering wings of emission lines formed
in line-driven hot star winds \citep{hillier1991eswingsmodel}. Complex substructures 
on a variety of scales are also produced in simulations of line-driven 
outflows in AGN \citep{PSK2000,PK04,progakurosawa2010,proga2014}.
Clumpy winds therefore offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

@@

There has been some success using simple
kinematic prescriptions for biconical disc winds to model 
AGN and quasar outflows \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}. 
H13 successfully produced a benchmark model for BALQSOs. However, the model
had two key drawbacks. First, an unrealistically low X-ray luminosity
was required in order to prevent over-ionization of the outflow.
Second, the model failed to produce the strong emission lines 
required at low inclinations in a unified model.
In this paper, we attempt to address these issues, and test the disc wind 
unification model using Monte Carlo radiative transfer (MCRT) and photoionization 
calculations. The paper is organised as follows.
In section 2, we describe some of the important photoionization 
and MCRT aspects of the code. In section 3, we outline the model, including 
a description of our clumping implementation. In section 4, we present the results 
from a clumped model. In section 5 we discuss our results, 
focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

We use the MCRT code \py\ to carry out our radiative transfer and photoionization
calculations in non-local thermodynamic equilibrium (non-LTE). 
The code has been used to model accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects \citep{simmacro2005}
and quasars (H13). \nocite{noebauer, M15, LK02}
These studies contain extensive detail on the code, 
so we only briefly describe the key elements of the global 
ionization calculation and other important aspects.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interact with either two-level `simple ions'
or full `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
Our scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H is treated as a macro-atom, because 
we expect recombination to be important
in determining its level populations and resultant line emission.

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by M15. Previously (LK02, H13, M15),
we adopted a modified Saha approach to calculate the ionization fractions
of simple-ions. We have now improved our code to explicitly solve the 
rate equations between ions in non-LTE. 
This dispenses with a number of small assumptions made in the modified Saha approach, 
is more numerically stable, 
and, in principle, allows the direct addition of extra physical 
processes such as Auger ionization.

%To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
We model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of $n$ bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $i$, or an exponential 
\begin{equation}
J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1}{J_{\nu,i}\sigma_j(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{DR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination, respectively.
We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

% We improve on this method by abandoning the modified Saha approach
% entirely, and instead computing the ionization state by 
% explicitly solving the rate equations in between ions in non-LTE. 
% Photoionization and recombination rates are calculated using MC estimators recorded
% during the photon propagation. Not only does this dispense with a number of
% small assumptions made in the modified Saha approach, it is also more numerically stable, 
% and in principle allows the direct addition of extra physical processes such as Auger ionization, which would have to be approximated if using the previous technique.
\subsection{Physical Processes}

@@

% MODEL DESCRIPTION

@@

Our kinematic prescription for a biconical disc wind model
follows \cite{SV93}, and is described further by
LK02, H13 and M15. A schematic is shown in figure~\ref{fig:cartoon},
with key aspects marked. The general biconical
geometry is similar to that invoked by \cite{MCGV95} and 
\cite{elvis2000} in order to explain the phenomenonology
of quasars and BALQSOs.

@@

% \subsection{A Benchmark Model for BALQSOs}

% \cite{higginbottom2013} presented a benchmark model for
% (BAL)QSOs...introduce key parameters.

@@

%Describe the photon sources in the model.

The accretion disc in our model is geometrically thin, but optically thick.
We thus treat the disc as an ensemble of blackbodies with a 
\cite{shakurasunyaev1973} effective temperature profile. 
The emergent SED is then determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are assumed to beopaque, meaning
that photons which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~r_G$, where 
$r_G = GM_{BH}/c^2$ is the gravitational radius.
For a $10^9~M_\odot$ black hole, this is equal to $8.8\times10^{14}~{\rm cm}$ 
or $\sim10^{-4}~{\rm pc}$.
